---
title: "candys"
author: "Juan Ignacio De Vicente"
date: "2022-10-25"
output: html_document
---





#### **A brief Introduction to the data set**

Who doesn't like candys? This data set is an overview about the consumers' taste during the years 2015,2016 and 2017. This survey was carry out in USA,so mainly it reflected the candy context in that country.However, there were a small proportion of participants from others countries. The value of each candy was marked as "JOY","MEH" and "DESPAIR".




#### **Assumptions**

My first assumption was before open the Task, I already knew I would have a lot of issues with the data in every stage of the process from star cleaning to answered the questions.

After the cleaning process - you never finish- I understood I should have a clear understandable of the data and the questions and a strategic before start tiding about columns/ variables (rename, keep or remove, class of data, number of columns per year), or rows/ observations ( leave or remove NA, )

I investigated the possibility to have all the variables in alphabetical order. I saw that in order to do that I should change the names during the cleaning process if I wanted to have few specific variables at the start of the table like "year" or "age"

At the start of the cleaning, I wanted to drop all NA values because I thought it would help me during the analysis, but later after read the questions I fell I could not do that, because it was like cheat. After the first question I realized, I should done it.  

Pivot. During the cleaning process is good examine the questions, just in case that is suitable/ necessary pivot columns/rows



#### **Cleaning steps.**

My first step was import and run all the possible libraries I would needed and import the raw files.

After import the raw data files, I thought to join them before to do any other operation. Fortunately,I was warned to avoid that idea. As I said  before, my lack of a clear tidy strategic made me loss and waste time.

I opened each year once I considered I was finished with the previous year. In each year I develop different ways to remove the columns I thought that they were unnecessary. I used janitor to clear names and I did manual rename. That option, take me to much time for the year 2017, later I learn that I was able to do it with a loop but I was unsure to do that.

I mutate one column to have one "2017".And i experimented issues with the age column and I tried to resolve. However, during the first question I understood I did not do it properly and I did again.
In addition, I used "pivot_longer" in order to answered  question 4




Every time I did any operation, I run the functions "view()" and "name()" to check if it was done what I wanted to do. I delete them from the "cleaning_script.R"





The last cleaning was done on Thursday. When I tried to resolve question 2, I saw that there was a problem with the data on the variable "age"and I tried to resolve. There were numbers with decimals. 

```{r, eval = FALSE}
candy_master %>% 
  mutate(age = str_replace(age,"\\..*", "")) %>% 
  mutate(age = as.numeric(age))

candy_master$age[candy_master$age > 120] <- 0
```
Also, in question 4 , it was needed to do a pivot_longer to convert all the candy  columns in observations and made easier the ratings.However, in this occasion , it was not modified the final file "candy_master".  

```{r, eval = FALSE}
  pivot_candy_master <- candy_master %>% 
  pivot_longer(cols = -c(1:3),names_to = "candy_question", values_to = "candy_rating")
```






#### **Answers**

Q1 What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)
```{r, eval = FALSE}

 total_rates <- candy_master %>% 
   summarise(across(.cols = -c(1:3), ~sum(is.na(.x)))) %>% 
   rowSums()
  
  total_rates #386685
```

Q2 What was the average age of people who are going out trick or treating?
```{r, eval = FALSE}


average_people_yes <- candy_master %>% 
  filter(trick_or_treat == "Yes") %>%
   group_by(trick_or_treat) %>% 
     summarise(mean_age = mean(age, na.rm = TRUE)) %>% 
      arrange(mean_age)
  

average_people_yes #34.92672
```

Q3 What was the average age of people who are not going trick or treating?
```{r, eval = FALSE}
average_people_no <- candy_master %>% 
  filter(trick_or_treat == "No") %>%
   group_by(trick_or_treat) %>% 
     summarise(mean_age = mean(age, na.rm = TRUE)) %>% 
      arrange(desc(mean_age))
  

average_people_no # 39.11304
```

Q4 For each of joy, despair and meh, which candy bar received the most of these ratings?
```{r, eval = FALSE}
pivot_candy_master <- candy_master %>% 
  pivot_longer(cols = -c(1:3),names_to = "candy_question", values_to = "candy_rating")

  pivot_candy_master %>% 
   filter(candy_rating == "JOY") %>% 
    group_by(candy_question) %>% 
     count() %>% 
      arrange(desc(n))
     
# Answer. Joy 3651 dark_chocolate_hershey
  
  pivot_candy_master %>% 
   filter(candy_rating == "DESPAIR") %>% 
    group_by(candy_question) %>% 
     count() %>% 
      arrange(desc(n))
  
#Answer. Despair 7288 Kale_smoothie
  
  pivot_candy_master %>% 
   filter(candy_rating == "MEH") %>% 
    group_by(candy_question) %>% 
     count() %>% 
      arrange(desc(n))
  
#Answer. Meh 1570 lollipops  

```


Q5 How many people rated Starburst as despair?
```{r, eval = FALSE}
starbust_despair <- candy_master %>% 
  count(sum(starburst == "DESPAIR"))


starbust_despair #9349 This is wrong
```

Q6 What was the most popular candy bar by this rating system for each gender in the data set ?
```{r, eval = FALSE}

```

Q7 What was the most popular candy bar in each year?
```{r, eval = FALSE}

```

Q8 What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?
```{r, eval = FALSE}

```








#### **Comments and conclusions you come across**