---
title: "R Notebook"
output: html_notebook
---

#mutate columns regex pivot all columns




```{r}
library(readxl)
library(tidyverse)
library(here)
library(dplyr)
library(janitor)


candy_2015 <- read_excel(here("raw_data/boing-boing-candy-2015.xlsx"))
candy_2016 <- read_excel(here("raw_data/boing-boing-candy-2016.xlsx"))
candy_2017 <- read_excel(here("raw_data/boing-boing-candy-2017.xlsx"))

names(candy_2015)
```

```{r}
# try to remove columns by starts with
clean_candy_2015 <- candy_2015 %>% select(-starts_with('P'))

names(clean_candy_2015)
```

```{r}
# Remove columns by range

clean_candy_2015 <- candy_2015 %>% select(c(1:96))

names(clean_candy_2015)
```

```{r}
#Janitor, clean names
janitor_clean_candy_2015 <- clean_candy_2015 %>% 
  clean_names()

view(janitor_clean_candy_2015)
```

```{r}
rename_janitor_clean_candy_2015 <- janitor_clean_candy_2015 %>% 
  rename("age" ="how_old_are_you", "year"= "timestamp","trick_or_treat" = "are_you_going_actually_going_trick_or_treating_yourself")

view(rename_janitor_clean_candy_2015)
names(rename_janitor_clean_candy_2015)
```

```{r}
#remove more colums

remove_col_rename_janitor_clean_candy_2015 <- rename_janitor_clean_candy_2015 %>% 
  select(c(1:5,8:11,13:15,19:22,24:25,27:30,32,35:37,39:84,86:92,96))


# remove_col_rename_janitor_clean_candy_2015 <- rename_janitor_clean_candy_2015 %>% 
#   select(6,7,12,16,17,18,23,26,31,33,34,38,85,93,94,95)


names(remove_col_rename_janitor_clean_candy_2015)
view(remove_col_rename_janitor_clean_candy_2015)
```


```{r}

#Change year

year_remove_col_rename_janitor_clean_candy_2015 <-  remove_col_rename_janitor_clean_candy_2015 %>% 
  mutate(year = coalesce("2015"))

view(year_remove_col_rename_janitor_clean_candy_2015)

```

```{r}
#Change age

age_year_remove_col_rename_janitor_clean_candy_2015 <- year_remove_col_rename_janitor_clean_candy_2015 %>% 
  mutate(age = str_replace(age,"\\..*", "")) %>% 
  mutate(age = as.numeric(age))
 
    # mutate(age = round(age,digit = 0))


view(age_year_remove_col_rename_janitor_clean_candy_2015)
```




```{r}
# # drop na . I did at the beginning, later I thought it was not a good idea
# janitor_clean_candy_2015 <- janitor_clean_candy_2015 %>% 
#   drop_na()
# view(janitor_clean_candy_2015)
```















```{r}
#Start cleaning candy_2016

names(candy_2016)
```
```{r}
#remove columns by range
clean_candy_2016 <- candy_2016 %>% select(-c(104:123))

names(clean_candy_2016)
```

```{r}
# Remove by column by index

clean_candy_2016 <- clean_candy_2016 %>% 
  select(-c(79,102))


clean_candy_2016 <- clean_candy_2016 %>% 
  select(-c(6))

names(clean_candy_2016)

```

```{r}
#janitor, clean names

janitor_clean_candy_2016 <- clean_candy_2016 %>% 
  clean_names()

view(janitor_clean_candy_2016)



rename_janitor_clean_candy_2016 <-janitor_clean_candy_2016 %>% 
  rename("gender" = "your_gender","age" = "how_old_are_you")


rename_janitor_clean_candy_2016
view(rename_janitor_clean_candy_2016)
names(rename_janitor_clean_candy_2016)
```




```{r}
#more cleaning, not properly done

rename_janitor_clean_candy_2016 <- rename_janitor_clean_candy_2016 %>% 
  select(-c(6,8,12,19,22,23,26,27,31,38,43,93,100))


rename_janitor_clean_candy_2016 <- rename_janitor_clean_candy_2016 %>% 
  rename("toblerone_something_or_other" = "tolberone_something_or_other")


rename2_janitor_clean_candy_2016 <- rename_janitor_clean_candy_2016 %>% 
  rename("country" = "which_country_do_you_live_in", "year"= "timestamp", "trick_or_treat"= "are_you_going_actually_going_trick_or_treating_yourself")




remove_col_rename_janitor_clean_candy_2016 <- rename2_janitor_clean_candy_2016 %>% 
  select(-c(6,9,15,16,17,18,19,22,25,28,32,57,76,81,87))

names(remove_col_rename_janitor_clean_candy_2016)

view(remove_col_rename_janitor_clean_candy_2016)

```

```{r}
#change timeslap 2016
```


```{r}
year_remove_col_rename_janitor_clean_candy_2016 <-  remove_col_rename_janitor_clean_candy_2016 %>% 
  mutate(year = coalesce("2016"))
```


```{r}
view(year_remove_col_rename_janitor_clean_candy_2016)
```



```{r}
#round age in 2016

age_year_remove_col_rename_janitor_clean_candy_2016 <- year_remove_col_rename_janitor_clean_candy_2016 %>% 
  mutate(age = as.numeric(age)) %>% 
  mutate(age = round(age,digit = 0))



view(age_year_remove_col_rename_janitor_clean_candy_2016)
```


```{r}
#more cleaning

#country 2016 remove_col_rename_janitor_clean_candy_2016

country_year_remove_col_rename_janitor_clean_candy_2016 <- age_year_remove_col_rename_janitor_clean_candy_2016 %>%
  mutate(country = recode(country,"us"= "USA", "england"= "UK","england" = "UK", "Unite states of America" = "USA", "England" = "UK", "uSA" = "USA", "france"=  "France", "croatia" = "Croatia", "USSA"= "USA","USA (I think but it's an election year so who can really tell)" = "USA" , "United States" = "USA","usa"= "USA","uk"= "UK","USA!" = "USA","U.S.A." = "USA","51.0" ="NA","canada"= "Canada","America"= "USA","USA! USA! USA!"= "USA","US"= "USA","United States of America" = "USA","united states"= "USA","Usa" ="USA","U.S."= "USA","Us" ="USA","US" = "USA","Units States" ="USA","belgium" = "Belgium","united states"= "USA","USA USA USA" ="USA", "the best one-usa" = "USA","españa"= "España","u.s."= "USA","United Kingdom"= "UK","hungary"= "Hungary", "united states of america" ="USA","United states" = "USA", "USA! USA!" = "USA"))




view(country_year_remove_col_rename_janitor_clean_candy_2016)
# No work, why
# country_rename_janitor_clean_candy_2016 <- remove_col_rename_janitor_clean_candy_2016 %>%
#   mutate(country = recode(country,c("us","usa","United States of America")= "USA"))

```



```{r}
# # Drop NA. Same as 2015, I keep them.
# janitor_clean_candy_2016 <- janitor_clean_candy_2016 %>% 
#   drop_na()
# 
# view(janitor_clean_candy_2016)
```











```{r}
#Start cleanig data.
names(candy_2017)
view(candy_2017)
```



```{r}

clean_candy_2017 <-candy_2017 %>% 
  select(-c(6,12,19,21,22,23,25,26,27,31,35,38,43,69,70,75,81,96,104,105,107,108,110:120))

names(clean_candy_2017)
#No more cleaning
# clean_candy_2017 <- clean_candy_2017 %>% 
#   select(-c(16,93:95))
# 
# names(clean_candy_2017)
# 
# clean_candy_2017 <- clean_candy_2017 %>% 
#   select(-c(6,14,18,24,82,89))
# 
# names(clean_candy_2017)
```

```{r}

#janitor

janitor_clean_candy_2017 <- clean_candy_2017 %>% 
  clean_names()

view(janitor_clean_candy_2017)
names(janitor_clean_candy_2017)
```



```{r}
#Manually, rename columns, I was not able to create a loop for remove "q1:9_"

rename_janitor_clean_candy_2017 <- janitor_clean_candy_2017 %>% 
  rename("gender" =  "q2_gender","age" = "q3_age", "country" = "q4_country","100_grand_bar" ="q6_100_grand_bar", "those_odd_marshmallow_circus_peanut_things" =  "q6_those_odd_marshmallow_circus_peanut_things","three_musketeers"="q6_three_musketeers","toblerone_something_or_other" = "q6_tolberone_something_or_other", "trail_mix"="q6_trail_mix" ,"twix" ="q6_twix","whatchamacallit_bars"= "q6_whatchamacallit_bars","york_peppermint_patties"="q6_york_peppermint_patties","anonymous_brown_globs_that_come_in_black_and_orange_wrappers_a_k_a_mary_janes"= "q6_anonymous_brown_globs_that_come_in_black_and_orange_wrappers_a_k_a_mary_janes",  "black_jacks" ="q6_black_jacks",
  "bonkers_the_candy" = "q6_bonkers_the_candy","bottle_caps"= "q6_bottle_caps","boxo_raisins" = "q6_boxo_raisins", "butterfinger" ="q6_butterfinger" ,    "cadbury_creme_eggs" = "q6_cadbury_creme_eggs", "candy_corn" = "q6_candy_corn", "caramellos"  = "q6_caramellos",  "chiclets"  ="q6_chiclets","dots" = "q6_dots","dove_bars" = "q6_dove_bars", "fuzzy_peaches" ="q6_fuzzy_peaches", "goo_goo_clusters" = "q6_goo_goo_clusters", "good_n_plenty" ="q6_good_n_plenty",  "gummy_bears_straight_up" = "q6_gummy_bears_straight_up" ,"hard_candy" ="q6_hard_candy","heath_bar" ="q6_heath_bar",  " hersheys_dark_chocolate"  = "q6_hersheys_dark_chocolate" ,"hershey_s_milk_chocolate"  ="q6_hershey_s_milk_chocolate" , "hersheys_kisses"  ="q6_hersheys_kisses",  "jolly_rancher_bad_flavor" = "q6_jolly_rancher_bad_flavor","jolly_ranchers_good_flavor" = "q6_jolly_ranchers_good_flavor" , "joy_joy_mit_iodine" ="q6_joy_joy_mit_iodine","junior_mints"="q6_junior_mints","senior_mints"="q6_senior_mints","kale_smoothie" ="q6_kale_smoothie", "kinder_happy_hippo" =  "q6_kinder_happy_hippo","kit_kat" ="q6_kit_kat" , "laffy_taffy" = "q6_laffy_taffy", "q6_lemon_heads"="q6_lemon_heads","licorice_not_black" = "q6_licorice_not_black",             "licorice_yes_black"="q6_licorice_yes_black" ,"lindt_truffle" = "q6_lindt_truffle", "lollipops" = "q6_lollipops" ,  "mars"= "q6_mars",  
"maynards" = "q6_maynards", "mike_and_ike"  ="q6_mike_and_ike" ,"milk_duds" ="q6_milk_duds","milky_way" ="q6_milky_way" ,"regular_m_ms" ="q6_regular_m_ms",   "peanut_m_m_s"="q6_peanut_m_m_s", "blue_m_ms"  ="q6_blue_m_ms", "red_m_ms" ="q6_red_m_ms","green_party_m_ms"= "q6_green_party_m_ms",    "independent_m_ms"="q6_independent_m_ms","mint_kisses" ="q6_mint_kisses" , "mint_juleps" ="q6_mint_juleps" , "mr_goodbar"= "q6_mr_goodbar" , "necco_wafers"= "q6_necco_wafers","nestle_crunch"  ="q6_nestle_crunch" ,"nown_laters" ="q6_nown_laters", "peeps"  = "q6_peeps",  "pencils"  ="q6_pencils" ,"pixy_stix" = "q6_pixy_stix", "reese_s_peanut_butter_cups" ="q6_reese_s_peanut_butter_cups", "reeses_pieces"  = "q6_reeses_pieces" ,"reggie_jackson_bar" = "q6_reggie_jackson_bar", "rolos"  = "q6_rolos" ,           "sandwich_sized_bags_filled_with_boo_berry_crunch" =  "q6_sandwich_sized_bags_filled_with_boo_berry_crunch" , "skittles" = "q6_skittles" ,                   "smarties_american"  ="q6_smarties_american" , "smarties_commonwealth" = "q6_smarties_commonwealth" ,"snickers" =  "q6_snickers" ,                
  "sourpatch_kids_i_e_abominations_of_nature" = "q6_sourpatch_kids_i_e_abominations_of_nature","spotted_dick" = "q6_spotted_dick", "starburst"  ="q6_starburst","sweet_tarts" = "q6_sweet_tarts" ,  "swedish_fish" = "q6_swedish_fish",  "tic_tac" = "q6_tic_tacs", "lemon_heads" ="q6_lemon_heads" )


# try to do a loop but I didnt know
#for (col in ncolumnas)) {
 # rename_janitor_clean_candy_2017[, col]
#}


rename_janitor_clean_candy_2017 <- rename_janitor_clean_candy_2017 %>% 
  rename("year" = "internal_id","trick_or_treat" = "q1_going_out","mary_janes" = "anonymous_brown_globs_that_come_in_black_and_orange_wrappers_a_k_a_mary_janes","any_full_sized_candy_bar" = "q6_any_full_sized_candy_bar" )

#No more remove columns
# # remove_col_rename_janitor_clean_candy_2017 <- rename_janitor_clean_candy_2017 %>% 
#   select(-c(15,52,59,67))

view(rename_janitor_clean_candy_2017)
names(rename_janitor_clean_candy_2017)

```


```{r}
#Changeinternal_id for year 2017


year_rename_janitor_clean_candy_2017 <- rename_janitor_clean_candy_2017 %>% 
  mutate(year = coalesce("2017"))

view(year_rename_janitor_clean_candy_2017)
```



```{r}
#country2017

country_rename_janitor_clean_candy_2017 <- year_rename_janitor_clean_candy_2017 %>%
  mutate(country = recode(country,"us"= "USA", "england"= "UK","england" = "UK", "Unite states of America" = "USA", "England" = "UK", "uSA" = "USA", "france"=  "France", "croatia" = "Croatia", "USSA"= "USA","USA (I think but it's an election year so who can really tell)" = "USA" , "United States" = "USA","usa"= "USA","uk"= "UK","USA!" = "USA","U.S.A." = "USA","51.0" ="NA","canada"= "Canada","America"= "USA","USA! USA! USA!"= "USA","US"= "USA","United States of America" = "USA","united states"= "USA","Usa" ="USA","U.S."= "USA","Us" ="USA","US" = "USA","Units States" ="USA","belgium" = "Belgium","united states"= "USA","USA USA USA" ="USA", "the best one-usa" = "USA","españa"= "España","u.s."= "USA","United Kingdom"= "UK","hungary"= "Hungary", "united states of america" ="USA","United states" = "USA", "USA! USA!" = "USA","United staes" = "USA","u.s.a." ="USA","USAUSAUSA"= "USA","america"= "America","finland"="Findland","unhinged states"= "Unhinged states", "US of A"= "USA","Unites States" = "USA","The United States of America"= "USA", "The United States" = "USA", "Unied States" = "USA","germany" ="Germany","U S" = "USA","U.K." ="UK","unite satates" = "USA", "cascadaia"= "Cascadia","North Carolina" ="USA","Unhinged states" = "USA","cascadia" ="Cascadia", "canada"= "Canada","USA? Hard to tell anymore.." ="USA","'merica" = "America","Pittsburgh" ="USA","United State" = "USA","United Sates"= "USA","I pretend to be from Canada, but I am really from the United States" = "USA","United Stated" ="USA","Scotland"= "UK","New Jersey" = "USA","CANADA"="Canada","United ststes" = "USA","United Statss" = "USA","enland" = "UK","united States" = "UK","United Statea" = "USA","united ststes" = "USA","	
USA USA USA!!!!"  = "USA"))

view(country_rename_janitor_clean_candy_2017)
```

```{r}
#Problems columns numeric. Change 2017 age to numeric

age_country_rename_janitor_clean_candy_2017 <- country_rename_janitor_clean_candy_2017 %>% 
  mutate(age = as.numeric(age))

```


```{r}
#joins the tables the master tables and save as csv
view(age_year_remove_col_rename_janitor_clean_candy_2015)
view(country_year_remove_col_rename_janitor_clean_candy_2016)
view(age_country_rename_janitor_clean_candy_2017)


candy_master <- bind_rows(age_year_remove_col_rename_janitor_clean_candy_2015,
country_year_remove_col_rename_janitor_clean_candy_2016,age_country_rename_janitor_clean_candy_2017)

view(candy_master)



write_csv(candy_master,"../clean_data/candy_master.csv")
```

1.4.2 Analysis questions


```{r}
#try to put in alphabetical order columns. But not interested
candy_masterAZ = candy_master [ , order(c(names(candy_master)))]
names (candy_masterAZ)
```


What is the total number of candy ratings given across the three years. (Number of candy ratings, not the number of raters. Don’t count missing values)


```{r}
#number of columns time rows no.na 
#With NA
total_rates <- candy_master %>% 
  select(-c(1:3)) %>% 
   summarise(total_rates = n())
 
# filter(is.na(-c(1:3))) %>%
total_rates <- candy_master %>% 
  select(-c(1:3))%>%
   filter(na.rm = TRUE)
   summarise(total_rates = n())

total_rates
?across

```




What was the average age of people who are going out trick or treating?


```{r}

average_people <-  candy_master %>%
  select(age,trick_or_treat) %>% 
   filter(trick_or_treat == "yes") %>%
    mean(trick_or_treat =="yes", na.rm = TRUE)

average_people
```







What was the average age of people who are not going trick or treating?


For each of joy, despair and meh, which candy bar received the most of these ratings?

How many people rated Starburst as despair?








For the next three questions, count despair as -1, joy as +1, and meh as 0.

What was the most popular candy bar by this rating system for each gender in the dataset ?
What was the most popular candy bar in each year?
What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?











